---
title: "Breaking News Story"
editor: visual
---

```{r}
#| output: false
library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
library(here)

```

```{r}

events_saved <- readRDS(here("events_saved.rds"))

```

*After loading the initial packages, I conducted some initial analysis to include in my story. First, I found the events that were most commonly held by candidates. To do this, I grouped by event type and then arranged them by their descending count.*

```{r}

events_saved %>% 
  group_by(event_type) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

```

*Next, I analyzed which states the candidates most commonly held their events in. First, i group by state and then arranged by descending count to develop my list.*

```{r}

events_saved %>% 
  group_by(state) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) 

```

*Now I created my function to be used for inline text in my story. Here, my function was intended to find the state that a given candidate held the highest number of his/her events in. For the purpose of my story, I used Joe Biden as the input of my function and found that the highest count of his events was in DC, which was inputted inline into my story.*

```{r}

cand_state <- function(cand){
  
   
  result <- events_saved %>% 
    filter(cand_lastname == cand) %>% 
    count(state, sort = TRUE) %>% 
    head(1) %>% 
    select(state)
  
  return(result)
}

cand_state("Biden")

```

*For my second inline function, I wanted to find which candidate held the most events in a given state. For the purpose of my story, I used Iowa as my input as it is the first primary caucus. After running the function, I found John Delaney had held the most events in Iowa up until this date. Delaney's name was subsequently added inline into my story as a result of the function.*

```{r}

state_cand <- function(frequented_state){
  
   
  result <- events_saved %>% 
    filter(state == frequented_state) %>% 
    count(cand_lastname, sort = TRUE) %>% 
    head(1) %>% 
    select(cand_lastname)
  
  return(result)
}

state_cand("IA")


```

# **Democratic Presidential candidates race to the finish line as event season heats up**

Event season has been heating up across the country as we near the beginning of the 2020 primaries. With the Iowa caucus set to kick off in February, candidates have been holding events across the country. The presumed front runner of the field, former Vice President Joe Biden, has been active since he announced his candidacy, particularly in `r cand_state("Biden")`, where he has most frequently hosted his events. In a crowded field, his presence in the nation's capital could be a key factor come time for the primary elections. While the field has been hosting events across the country, there has been a particular emphasis on the early primary states such as Iowa and New Hampshire, which have hosted over three-hundred events combined. Iowa has been a particular favorite for Maryland Rep. John Delaney, as `r state_cand("IA")` has held the most events in the state. Events have primarily centered around speeches from individual candidates; however, the field has also hosted numerous town halls, meet-and-greets, and cattle calls across the country. As the primaries loom, the country will continue to watch eagerly to see who, out of the vast pool of candidates vying for the nomination, can remain on the campaign trail. Yet it is nearly a guarantee that we can only expect the amount of events to drastically pick up as the clock continues to tick closer to the Iowa Caucus.
