---
title: "Voting Walkthrough"
author: "Julian Gorman"
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true
---

*Here, I loaded in the packages and data that we'll use to provide analysis.*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

```

```{r}

counties_demographics <- readRDS("counties_demographics.rds")
counties_elexresults <- readRDS("counties_elexresults.rds")
counties_rural_urban_suburban <- readRDS("counties_rural_urban_suburban.rds")


```

1)  *These are the mean percentage of the votes Republicans received in VA by county type*. *County types are arranged in terms of density, with "large central" being the most urban county types, while "small - noncore" represent the most rural.*

```{r}


counties_rural_urban_suburban %>% 
filter(state_code == "VA") %>% 
inner_join(counties_elexresults, counties_rural_urban_suburban, by = "geoid") %>%
    group_by(rural_urban) %>%
  summarise_at(vars(pct_r_2020), list(pct_r = mean))


```

2)  *This is the number of counties by type that each party won in Virginia. As we can see in the results, Democrats in Virginia heavily outperformed Republicans in larger counties, while Republicans performed better in less dense areas.*

```{r}

counties_rural_urban_suburban %>% 
filter(state_code == "VA") %>% 
inner_join(counties_elexresults, counties_rural_urban_suburban, by = "geoid") %>% group_by(rural_urban) %>% 
  count(winner_2020) 

```

3)  *This is the mean percentage of votes in Florida Republicans won in counties either above and below the national average of college graduates. The results showed that Republicans won a higher percentage in Florida counties that were below the national average of college graduates.*

```{r}

 counties_demographics %>% 
  filter(state_code == "FL") %>% 
  inner_join(counties_elexresults, counties_demographics, by = "pct_r_2020") %>%
  group_by(pct_ed_college_all_abovebelow_natl) %>% 
  summarise_at(vars(pct_r_2020), list(pct_r = mean)) 

  

```

4)  *This is the mean percentage of votes Republicans won in all U.S. counties sorted by their relation to the national average of college graduates. Similar to our analysis of just Florida, the results on a national scale show that Republicans won a much higher percentage in U.S. counties below the national average of college graduates.*

```{r}

  inner_join(counties_elexresults, counties_demographics, by = "pct_r_2020") %>%
  group_by(pct_ed_college_all_abovebelow_natl) %>% 
  summarise_at(vars(pct_r_2020), list(pct_r = mean)) 


```

5)  *Here I found the mean percentage of votes for Trump in the 2020 election in counties by percentage of foreign born people in comparison to the national average. In 2020, the results show that Trump recorded over ten points higher in counties that had a percentage of foreign people lower than the national average.*

```{r}

  inner_join(counties_elexresults, counties_demographics, by = "pct_r_2020") %>% 
  group_by(pct_born_foreign_abovebelow_natl) %>% 
  summarise_at(vars(pct_r_2020), list(pct_r = mean)) 


```

6)  *Here I found the count of counties won by Democrats and Republicans grouped by the nonwhite percentage in comparison to the national average. As we can see, Republicans heavily outperformed Democrats in counties that had a nonwhite percentage below the national average.*

```{r}

inner_join(counties_elexresults, counties_demographics, by = "geoid") %>% group_by(pct_race_nonwhite_abovebelow_natl) %>% 
  count(winner_2020) 



```

7)  *Here I joined all three major datasets together into one. This allowed us to incorporate all of the major relevant data measures into one sheet.*

```{r}

inner_join(counties_demographics, counties_elexresults, counties_rural_urban_suburban, by = "geoid")




```
